---
title: "Mapping NEON Tick and Tick Pathogen Data Survey Monitoring Data to Humboldt Extension"
subtitle: "See: [Guide for publishing survey & monitoring data (GBIF Draft)](https://docs.gbif-test.org/guide-publishing-survey-data/en/)"
author: "Chandra Earl"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

<https://github.com/sunray1/NEONTickstoHumboldt>

The following code will download tick and tick pathogen data collected by the National Ecological Observatory Network (NEON). The goal is to remap the data to Darwin-Core using the Humboldt Extension. This will include vouchered specimens (Occurrence) and event ecological data (Humboldt).

# Load libraries

```{r message=FALSE, warning=FALSE}
library(neonUtilities)
library(dplyr)
```

# Download files

[NEON 2025a] NEON (National Ecological Observatory Network) [1]. 2025. NEON Tick pathogen status (DP1.10092.001), RELEASE-2025. https://doi.org/10.48443/8nhe-cp13. Dataset accessed from https://data.neonscience.org/data-products/DP1.10092.001/RELEASE-2025 on 2025-09-25.

[NEON 2025b] NEON (National Ecological Observatory Network) [2]. 2025. NEON Ticks sampled using drag cloths (DP1.10093.001), RELEASE-2025. https://doi.org/10.48443/6zpz-5z19. Dataset accessed from https://data.neonscience.org/data-products/DP1.10093.001/RELEASE-2025 on 2025-09-25.

I have a NEON token to download data, though I've saved the data as an Rdata file so users don't have to redownload.

## Get Tick and Tick Pathogen data

```{r}
if (!dir.exists("data")) {
  dir.create("data")
}

if (!dir.exists("outputs")) {
  dir.create("outputs")
}

if (file.exists("data/tick.pathogen.RData")) {
  load("data/tick.pathogen.RData")
} else {
  
  tick.pathogen <- loadByProduct(dpID="DP1.10092.001", 
                                 package = "basic",
                                 release = "RELEASE-2025",
                                 token = Sys.getenv("NEON_TOKEN"),
                                 check.size = F)
  
  save(tick.pathogen, file = "data/tick.pathogen.RData")
}

if (file.exists("data/tick.occurrence.RData")) {
  load("data/tick.occurrence.RData")
} else {
  
  tick.occurrence <- loadByProduct(dpID="DP1.10093.001", 
                                   package = "basic",
                                   release = "RELEASE-2025",
                                   token = Sys.getenv("NEON_TOKEN"),
                                   check.size = F)
  
  save(tick.occurrence, file = "data/tick.occurrence.RData")
}

```

# Map Data

## Map Event Data

Some event data is not included in the NEON data download (data for the project, domain and site). 

All fields from the [Survey event table template](https://github.com/gbif/doc-guide-publishing-survey-data/tree/1.1_postCommReview/data/event_template_wHE_event-table.csv) are included for completeness, even if some are not applicable to this dataset. 

### Project Event

```{r}
NEON_project_event <- data.frame(parentEventID = NA, #Q: should I add namespace to these terms?
                                  eventID = "NEON",
                                  siteNestingDescription = "46 terrestrial sites each with at least six 40×40 m plots designated for tick sampling (plots may be decommissioned and reassigned as necessary, but six plots are required for sampling) located across 19 ecoclimatic domains",
                                  siteCount = length(unique(tick.occurrence$tck_fielddata$plotID)),
                                  fieldNumber = NA,
                                  verbatimSiteNames = paste(sort(unique(tick.occurrence$tck_fielddata$plotID)), collapse = " | "),
                                  habitat = NA,
                                  verbatimSiteDescriptions = NA,
                                  reportedWeather = NA,
                                  reportedExtremeConditions = NA,
                                  locationID = NA,
                                  countryCode = "US",
                                  decimalLatitude = NA,
                                  decimalLongitude = NA,
                                  coordinateUncertaintyInMeters = NA,
                                  geodeticDatum = NA,
                                  locality = NA,
                                  totalAreaSampledValue = length(unique(tick.occurrence$tck_fielddata$plotID)) * 40 * 4, #summed value of plot perimeters
                                  totalAreaSampledUnit = "m",
                                  geospatialScopeAreaValue = 9428288, #full geographic extent of USA-NEON boundaries
                                  geospatialScopeAreaUnit = "km²",
                                  sampleSizeValue = NA,
                                  sampleSizeUnit = NA,
                                  footprintWKT = NA,
                                  footprintSRS = NA,
                                  isVegetationCoverReported = "false",
                                  eventDate = paste(
                                    min(tick.occurrence$tck_fielddata$collectDate, na.rm = TRUE),
                                    max(tick.occurrence$tck_fielddata$collectDate, na.rm = TRUE),
                                    sep = "/"
                                  ),
                                  eventTime = NA,
                                  eventDurationValue = NA,
                                  eventDurationUnit = NA,
                                  eventType = "project",
                                  inventoryTypes = NA, #Q: I guess we could consider this an inventory and fill out this and the following, though the guide says only if eventType = Inventory
                                  compilationTypes = NA,
                                  compilationSourceTypes = NA,
                                  samplingProtocol = "Drag sampling | Flagging",
                                  protocolNames = "Drag sampling | Flagging",
                                  protocolDescriptions = "Ticks are sampled at six plots per site every three to six weeks, depending on prior detections. Sampling follows the 160 m perimeter of each 40 × 40 m plot, using drag cloths (or flagging in dense vegetation). Ticks from all life stages are collected every few meters and preserved in ethanol, with bouts scheduled between green-up and dormancy. Samples are sent to external labs where they are sorted, counted and a subset of identified nymphs are tested for pathogens.",
                                  protocolReferences = "Paull, S. 2022. TOS Protocol and Procedure: TCK – Tick and Tick-Borne Pathogen Sampling. NEON.DOC.014045. NEON (National Ecological Observatory Network). | Laboratory of Medical Zoology (LMZ). 2023. NEON Tick Pathogen Testing SOP, V4.01. University of Massachusetts, Amherst. | Beati, L. 2021. Tick Identification Instructions, USNTC Standard Operating Procedure (SOP). V3. Georgia Southern University, US National Tick Collection (USNTC).",
                                  isAbsenceReported = "true", #Q: Should this be reported true at this level or only at the event visit level?
                                  absentTaxa = NA,
                                  isAbundanceReported = "true",
                                  isAbundanceCapReported = "false",
                                  abundanceCap = NA,
                                  hasMaterialSamples = "true", #Q: Since we're not explicitly designating these as material samples, should this be false?
                                  materialSampleTypes = "wholeOrganism",
                                  hasVouchers = "true",
                                  voucherInstitutions = "US National Tick Collection, Georgia Southern University",
                                  isLeastSpecificTargetCategoryQuantityInclusive = "false",
                                  dataGeneralizations = NA,
                                  informationWithheld = NA, #TODO: double check this - I don't think any tick taxa are withheld but I'm not 100% certain
                                  fieldNotes = NA,
                                  eventRemarks = NA,
                                  verbatimTargetScope = "ticks",
                                  targetTaxonomicScope = "?Ixodidae", #Q: Since we have pathogens in occurrences, I'm not sure how to designate those here - the ones tested don't have a mca - most are Bacteria but one is a piroplasm
                                  excludedTaxonomicScope = NA,
                                  isTaxonomicScopeFullyReported = "true",
                                  taxonCompletenessReported = "reportedComplete", #Q: I'm not 100% sure I did these two correctly
                                  taxonCompletenessProtocols = "Tick drag/flag methods are standardized to comprehensively detect all life stages of Ixodidae during the active season.", #Q: I'm not 100% sure I did these two correctly
                                  hasNonTargetTaxa = "false",
                                  areNonTargetTaxaFullyReported = NA,
                                  nonTargetTaxa = NA,
                                  identifiedBy = paste(unique(tick.occurrence$tck_taxonomyProcessed$laboratoryName), collapse = " | "), #Q: I could change this to paste(unique(tick.occurrence$tck_taxonomyProcessed$identifiedBy), collapse = " | ") and get the actual list of people who did the identifications, though this is not clean
                                  targetLifeStageScope = "adult | nymph | larvae",
                                  excludedLifeStageScope = "egg",
                                  isLifeStageScopeFullyReported = "true",
                                  targetDegreeOfEstablishmentScope = NA,
                                  excludedDegreeOfEstablishmentScope = NA,
                                  isDegreeOfEstablishmentScopeFullyReported = NA,
                                  targetGrowthFormScope = NA,
                                  excludedGrowthFormScope = NA,
                                  isGrowthFormScopeFullyReported = NA,
                                  hasNonTargetOrganisms = NA,
                                  targetHabitatScope = NA,
                                  excludedHabitatScope = NA,
                                  samplingEffort = "6 plots per bout, 160 m perimeter sampled per plot",
                                  isSamplingEffortReported = "true",
                                  samplingEffortProtocol = "Ticks are sampled at six plots per site, with bouts every 3 or 6 weeks depending on intensity. Each bout consists of dragging (or flagging if needed) along the full 160 m perimeter of each 40 × 40 m plot. Ticks of all life stages are collected at intervals and preserved in ethanol.",
                                  samplingEffortValue = "6, 160",
                                  samplingEffortUnit = "plots per bout, meters per plot circuit",
                                  samplingPerformedBy = "NEON Field Staff" #Q: I could change this to paste(unique(tick.occurrence$tck_fielddata$measuredBy), collapse = " | ") and get the list of actual people, though this has orcid and emails and is quite messy
                                )
```

### Domain Events

```{r}
NEON_domain_metadata <- read.csv("data/neon_domain_METADATA.csv", stringsAsFactors = FALSE)

NEON_domain_list <- unique(tick.occurrence$tck_fielddata[, c("domainID")])

#index
domain_idx <- match(NEON_domain_list, NEON_domain_metadata$domain_id)

added_domain_data <- data.frame(parentEventID = "NEON",
                                eventID = NEON_domain_metadata$domain_id[domain_idx],
                                siteNestingDescription = "Each domain contains 1-3 terrestrial field sites, each with at least six 40×40 m plots designated for tick sampling (plots may be decommissioned and reassigned as necessary, but six plots are required for sampling)",
                                siteCount = tapply(tick.occurrence$tck_fielddata$plotID,    #Get number of sampled plots per domain
                                                   tick.occurrence$tck_fielddata$domainID,
                                                   function(x) length(unique(x)))[NEON_domain_list],
                                fieldNumber = NA,
                                verbatimSiteNames = tapply(tick.occurrence$tck_fielddata$plotID,     #Get list of sampled plots at each domain
                                                           tick.occurrence$tck_fielddata$domainID,
                                                           function(x) paste(sort(unique(x)), collapse = " | ")
                                                         )[NEON_domain_list],
                                habitat = NA,
                                verbatimSiteDescriptions = NA,
                                reportedWeather = NA,
                                reportedExtremeConditions = NA,
                                locationID = NEON_domain_metadata$domain_id[domain_idx],
                                countryCode = "US",
                                decimalLatitude = NEON_domain_metadata$decimalLatitude[domain_idx],
                                decimalLongitude = NEON_domain_metadata$decimalLongitude[domain_idx],
                                coordinateUncertaintyInMeters = NEON_domain_metadata$coordinateUncertaintyinMeters[domain_idx],
                                geodeticDatum = "WGS84",
                                locality = paste(NEON_domain_metadata$domain_name[domain_idx], " (", NEON_domain_metadata$domain_id[domain_idx], ")", sep = ""),
                                totalAreaSampledValue = tapply(tick.occurrence$tck_fielddata$plotID, #summed value of surveyed plot perimeters in domain
                                                               tick.occurrence$tck_fielddata$domainID,
                                                               function(x) length(unique(x)))[NEON_domain_list] * 40 * 4,
                                totalAreaSampledUnit = "m",
                                geospatialScopeAreaValue = NEON_domain_metadata$sq_km[domain_idx],
                                geospatialScopeAreaUnit = "km²",
                                sampleSizeValue = NA,
                                sampleSizeUnit = NA,
                                footprintWKT = NA, #Q: I'm not going to bother with these since they'll be huge
                                footprintSRS = NA,
                                isVegetationCoverReported = "false",
                                eventDate = tapply(
                                  tick.occurrence$tck_fielddata$collectDate,
                                  tick.occurrence$tck_fielddata$domainID,
                                  function(x) paste(min(x, na.rm = TRUE), max(x, na.rm = TRUE), sep = "/")
                                )[NEON_domain_list],
                                eventTime = NA,
                                eventDurationValue = NA,
                                eventDurationUnit = NA,
                                eventType = "domain",
                                inventoryTypes = NA,
                                compilationTypes = NA,
                                compilationSourceTypes = NA,
                                samplingProtocol = "Drag sampling | Flagging", #Sampling protocol terms should be populated for every Event regardless of hierarchical level Q: I took this to mean all the terms described in 5.5. Methodology or sampling protocol (from here up to eventRemarks), or is this just for the four protocol fields?
                                protocolNames = "Drag sampling | Flagging",
                                protocolDescriptions = "Ticks are sampled at six plots per site every three to six weeks, depending on prior detections. Sampling follows the 160 m perimeter of each 40 × 40 m plot, using drag cloths (or flagging in dense vegetation). Ticks from all life stages are collected every few meters and preserved in ethanol, with bouts scheduled between green-up and dormancy. Samples are sent to external labs where they are sorted, counted and a subset of identified nymphs are tested for pathogens.",
                                protocolReferences = "Paull, S. 2022. TOS Protocol and Procedure: TCK – Tick and Tick-Borne Pathogen Sampling. NEON.DOC.014045. NEON (National Ecological Observatory Network). | Laboratory of Medical Zoology (LMZ). 2023. NEON Tick Pathogen Testing SOP, V4.01. University of Massachusetts, Amherst. | Beati, L. 2021. Tick Identification Instructions, USNTC Standard Operating Procedure (SOP). V3. Georgia Southern University, US National Tick Collection (USNTC).",
                                isAbsenceReported = "true", #Q: Should this be reported true at this level or only at the event visit level?
                                absentTaxa = NA,
                                isAbundanceReported = "true",
                                isAbundanceCapReported = "false",
                                abundanceCap = NA,
                                hasMaterialSamples = "true", #Q: Since we're not explicitly designating these as material samples, should this be false?
                                materialSampleTypes = "wholeOrganism",
                                hasVouchers = "true",
                                voucherInstitutions = "US National Tick Collection, Georgia Southern University",
                                isLeastSpecificTargetCategoryQuantityInclusive = "false",
                                dataGeneralizations = NA,
                                informationWithheld = NA, #TODO: double check this - I don't think any tick taxa are withheld but I'm not 100% certain
                                fieldNotes = NA,
                                eventRemarks = NA,
                                verbatimTargetScope = "ticks", # Recommended best practice is to populate scope terms every Event to which they apply.
                                targetTaxonomicScope = "?Ixodidae", #Q: Since we have pathogens in occurrences, I'm not sure how to designate those here - the ones tested don't have a mca - most are Bacteria but one is a piroplasm
                                excludedTaxonomicScope = NA,
                                isTaxonomicScopeFullyReported = "true",
                                taxonCompletenessReported = "reportedComplete", #Q: I'm not 100% sure I did these two correctly
                                taxonCompletenessProtocols = "Tick drag/flag methods are standardized to comprehensively detect all life stages of Ixodidae during the active season.", #Q: I'm not 100% sure I did these two correctly
                                hasNonTargetTaxa = "false",
                                areNonTargetTaxaFullyReported = NA,
                                nonTargetTaxa = NA,
                                identifiedBy = tick.occurrence$tck_taxonomyProcessed %>% 
                                  left_join(tick.occurrence$tck_fielddata %>% select(sampleID, domainID),
                                            by = "sampleID") %>%
                                  group_by(domainID) %>%
                                  summarise(identifiedBy = paste(sort(unique(laboratoryName)), collapse = " | "),
                                            .groups = "drop") %>%
                                  right_join(data.frame(domainID = NEON_domain_list), by = "domainID") %>%
                                  arrange(match(domainID, NEON_domain_list)) %>%
                                  mutate(identifiedBy = ifelse(is.na(identifiedBy), NA, identifiedBy)) %>% #Q: Some domains/sites have had no ticks found, so identifiedBy is NA - should this be changed to something more descriptive, like 'No ticks found'?
                                  pull(identifiedBy), #Q: I could change this to get the actual list of people who did the identifications, though this is not clean
                                targetLifeStageScope = "adult | nymph | larvae",
                                excludedLifeStageScope = "egg",
                                isLifeStageScopeFullyReported = "true",
                                targetDegreeOfEstablishmentScope = NA,
                                excludedDegreeOfEstablishmentScope = NA,
                                isDegreeOfEstablishmentScopeFullyReported = NA,
                                targetGrowthFormScope = NA,
                                excludedGrowthFormScope = NA,
                                isGrowthFormScopeFullyReported = NA,
                                hasNonTargetOrganisms = NA,
                                targetHabitatScope = NA,
                                excludedHabitatScope = NA,
                                samplingEffort = "6 plots per bout, 160 m perimeter sampled per plot",
                                isSamplingEffortReported = "true",
                                samplingEffortProtocol = "Ticks are sampled at six plots per site, with bouts every 3 or 6 weeks depending on intensity. Each bout consists of dragging (or flagging if needed) along the full 160 m perimeter of each 40 × 40 m plot. Ticks of all life stages are collected at intervals and preserved in ethanol.",
                                samplingEffortValue = "6, 160",
                                samplingEffortUnit = "plots per bout, meters per plot circuit",
                                samplingPerformedBy = "NEON Field Staff" #Q: I could change this to paste(unique(tick.occurrence$tck_fielddata$measuredBy), collapse = " | ") and get the list of actual people, though this has orcid and emails and is quite messy
                              )
```


### Site Events

```{r}
NEON_site_metadata <- read.csv(
  "https://www.neonscience.org/field-sites/exports/NEON_Field_Site_Metadata_20250924",
  stringsAsFactors = FALSE
)

NEON_site_list <- unique(tick.occurrence$tck_fielddata[, c("siteID")])

#indexes
site_idx <- match(NEON_site_list, NEON_site_metadata$site_id)

added_site_data <- data.frame(parentEventID = NEON_site_metadata$domain_id[site_idx],
                              eventID = NEON_site_metadata$site_id[site_idx],
                              siteNestingDescription = "Each site contains at least six 40×40 m plots designated for tick sampling (plots may be decommissioned and reassigned as necessary, but six plots are required for sampling)",
                              siteCount = tapply(tick.occurrence$tck_fielddata$plotID,    #Get number of sampled plots per site
                                                 tick.occurrence$tck_fielddata$siteID,
                                                 function(x) length(unique(x)))[NEON_site_list],
                              fieldNumber = NA,
                              verbatimSiteNames = tapply(tick.occurrence$tck_fielddata$plotID,     #Get list of sampled plots at each site
                                                         tick.occurrence$tck_fielddata$siteID,
                                                         function(x) paste(sort(unique(x)), collapse = " | ")
                                                       )[NEON_site_list],
                              habitat = gsub("\\|", " | ", NEON_site_metadata$dominant_nlcd_classes[site_idx]),
                              verbatimSiteDescriptions = NEON_site_metadata$site_type[site_idx],
                              reportedWeather = NA,
                              reportedExtremeConditions = NA,
                              locationID = NEON_site_metadata$site_id[site_idx],
                              countryCode = "US",
                              county = NEON_site_metadata$site_county[site_idx],
                              stateProvince = NEON_site_metadata$site_state[site_idx],
                              decimalLatitude = NEON_site_metadata$latitude[site_idx],
                              decimalLongitude = NEON_site_metadata$longitude[site_idx],
                              coordinateUncertaintyInMeters = NA, #Q: I can attempt to get uncertainty for each site but we don't have it handy and I'd have to calculate it by hand
                              geodeticDatum = "WGS84",
                              locality = paste(NEON_site_metadata$site_name[site_idx], " (", NEON_site_metadata$site_id[site_idx], ")", sep = ""),
                              minimumElevationInMeters = NEON_site_metadata$minimum_elevation_m[site_idx],
                              maximumElevationInMeters = NEON_site_metadata$maximum_elevation_m[site_idx],
                              verbatimElevation = paste(NEON_site_metadata$mean_evelation_m[site_idx], "m", sep=""),
                              totalAreaSampledValue = tapply(tick.occurrence$tck_fielddata$plotID, #summed value of surveyed plot perimeters in site
                                                             tick.occurrence$tck_fielddata$siteID,
                                                             function(x) length(unique(x)))[NEON_site_list] * 40 * 4,
                              totalAreaSampledUnit = "m",
                              geospatialScopeAreaValue = NEON_site_metadata$terrestrial_sampling_boundary_size_km2[site_idx],
                              geospatialScopeAreaUnit = "km²",
                              sampleSizeValue = NA,
                              sampleSizeUnit = NA,
                              footprintWKT = NA, #Q: I'm not going to bother with these since they'll be huge
                              footprintSRS = NA,
                              isVegetationCoverReported = "false",
                              eventDate = tapply(
                                tick.occurrence$tck_fielddata$collectDate,
                                tick.occurrence$tck_fielddata$siteID,
                                function(x) paste(min(x, na.rm = TRUE), max(x, na.rm = TRUE), sep = "/")
                              )[NEON_site_list],
                              eventTime = NA,
                              eventDurationValue = NA,
                              eventDurationUnit = NA,
                              eventType = "site", 
                              inventoryTypes = NA,
                              compilationTypes = NA,
                              compilationSourceTypes = NA,
                              samplingProtocol = "Drag sampling | Flagging", #Sampling protocol terms should be populated for every Event regardless of hierarchical level Q: I took this to mean all the terms described in 5.5. Methodology or sampling protocol (from here up to eventRemarks), or is this just for the four protocol fields?
                              protocolNames = "Drag sampling | Flagging",
                              protocolDescriptions = "Ticks are sampled at six plots per site every three to six weeks, depending on prior detections. Sampling follows the 160 m perimeter of each 40 × 40 m plot, using drag cloths (or flagging in dense vegetation). Ticks from all life stages are collected every few meters and preserved in ethanol, with bouts scheduled between green-up and dormancy. Samples are sent to external labs where they are sorted, counted and a subset of identified nymphs are tested for pathogens.",
                              protocolReferences = "Paull, S. 2022. TOS Protocol and Procedure: TCK – Tick and Tick-Borne Pathogen Sampling. NEON.DOC.014045. NEON (National Ecological Observatory Network). | Laboratory of Medical Zoology (LMZ). 2023. NEON Tick Pathogen Testing SOP, V4.01. University of Massachusetts, Amherst. | Beati, L. 2021. Tick Identification Instructions, USNTC Standard Operating Procedure (SOP). V3. Georgia Southern University, US National Tick Collection (USNTC).",
                              isAbsenceReported = "true", #Q: Should this be reported true at this level or only at the event visit level?
                              absentTaxa = NA,
                              isAbundanceReported = "true",
                              isAbundanceCapReported = "false",
                              abundanceCap = NA,
                              hasMaterialSamples = "true", #Q: Since we're not explicitly designating these as material samples, should this be false?
                              materialSampleTypes = "wholeOrganism",
                              hasVouchers = "true",
                              voucherInstitutions = "US National Tick Collection, Georgia Southern University",
                              isLeastSpecificTargetCategoryQuantityInclusive = "false",
                              dataGeneralizations = NA,
                              informationWithheld = NA, #TODO: double check this - I don't think any tick taxa are withheld but I'm not 100% certain
                              fieldNotes = NA,
                              eventRemarks = NA,
                              verbatimTargetScope = "ticks", # Recommended best practice is to populate scope terms every Event to which they apply.
                              targetTaxonomicScope = "?Ixodidae", #Q: Since we have pathogens in occurrences, I'm not sure how to designate those here - the ones tested don't have a mca - most are Bacteria but one is a piroplasm
                              excludedTaxonomicScope = NA,
                              isTaxonomicScopeFullyReported = "true",
                              taxonCompletenessReported = "reportedComplete", #Q: I'm not 100% sure I did these two correctly
                              taxonCompletenessProtocols = "Tick drag/flag methods are standardized to comprehensively detect all life stages of Ixodidae during the active season.", #Q: I'm not 100% sure I did these two correctly
                              hasNonTargetTaxa = "false",
                              areNonTargetTaxaFullyReported = NA,
                              nonTargetTaxa = NA,
                              identifiedBy = tick.occurrence$tck_taxonomyProcessed %>% 
                                left_join(tick.occurrence$tck_fielddata %>% select(sampleID, siteID),
                                          by = "sampleID") %>%
                                group_by(siteID) %>%
                                summarise(identifiedBy = paste(sort(unique(laboratoryName)), collapse = " | "),
                                          .groups = "drop") %>%
                                right_join(data.frame(siteID = NEON_site_list), by = "siteID") %>%
                                arrange(match(siteID, NEON_site_list)) %>%
                                mutate(identifiedBy = ifelse(is.na(identifiedBy), NA, identifiedBy)) %>% #Q: Some domains/sites have had no ticks found, so identifiedBy is NA - should this be changed to something more descriptive, like 'No ticks found'?
                                pull(identifiedBy), #Q: I could change this to get the actual list of people who did the identifications, though this is not clean
                              targetLifeStageScope = "adult | nymph | larvae",
                              excludedLifeStageScope = "egg",
                              isLifeStageScopeFullyReported = "true",
                              targetDegreeOfEstablishmentScope = NA,
                              excludedDegreeOfEstablishmentScope = NA,
                              isDegreeOfEstablishmentScopeFullyReported = NA,
                              targetGrowthFormScope = NA,
                              excludedGrowthFormScope = NA,
                              isGrowthFormScopeFullyReported = NA,
                              hasNonTargetOrganisms = NA,
                              targetHabitatScope = NA,
                              excludedHabitatScope = NA,
                              samplingEffort = "6 plots per bout, 160 m perimeter sampled per plot",
                              isSamplingEffortReported = "true",
                              samplingEffortProtocol = "Ticks are sampled at six plots per site, with bouts every 3 or 6 weeks depending on intensity. Each bout consists of dragging (or flagging if needed) along the full 160 m perimeter of each 40 × 40 m plot. Ticks of all life stages are collected at intervals and preserved in ethanol.",
                              samplingEffortValue = "6, 160",
                              samplingEffortUnit = "plots per bout, meters per plot circuit",
                              samplingPerformedBy = "NEON Field Staff" #Q: I could change this to paste(unique(tick.occurrence$tck_fielddata$measuredBy), collapse = " | ") and get the list of actual people, though this has orcid and emails and is quite messy
                            )

```

### Plot Events

```{r}
NEON_plot_metadata <- unique(tick.occurrence$tck_fielddata[, c("namedLocation",
                                                               "siteID",
                                                               "plotID",
                                                               "plotType",
                                                               "nlcdClass",
                                                               "decimalLatitude",
                                                               "decimalLongitude",
                                                               "geodeticDatum",
                                                               "coordinateUncertainty",
                                                               "elevation",
                                                               "elevationUncertainty")])

NEON_plot_metadata <- NEON_plot_metadata %>% #There's an issue where plot JERC_081 has two elevations, causing it to be duplicated, so we'll just remove the first one
  distinct(namedLocation, plotID, .keep_all = TRUE)

added_plot_data <- data.frame(parentEventID = NEON_plot_metadata$siteID,
                              eventID = NEON_plot_metadata$plotID,
                              siteNestingDescription = "One 40×40 m plot designated for tick sampling",
                              siteCount = 1,
                              fieldNumber = NA,
                              verbatimSiteNames = NEON_plot_metadata$plotID,
                              habitat = NEON_plot_metadata$nlcdClass,
                              verbatimSiteDescriptions = NA,
                              reportedWeather = NA,
                              reportedExtremeConditions = NA,
                              locationID = NEON_plot_metadata$plotID,
                              countryCode = "US",
                              decimalLatitude = NEON_plot_metadata$decimalLatitude,
                              decimalLongitude = NEON_plot_metadata$decimalLongitude,
                              coordinateUncertaintyInMeters = NEON_plot_metadata$coordinateUncertainty,
                              geodeticDatum = "WGS84",
                              locality = NEON_plot_metadata$plotID,
                              minimumElevationinMeters = NEON_plot_metadata$elevation - NEON_plot_metadata$elevationUncertainty,
                              maximumElevationinMeters = NEON_plot_metadata$elevation + NEON_plot_metadata$elevationUncertainty,
                              verbatimElevation = paste(NEON_plot_metadata$elevation, "m", sep=""),
                              totalAreaSampledValue = 40 * 4,
                              totalAreaSampledUnit = "m",
                              geospatialScopeAreaValue = 40 * 4,
                              geospatialScopeAreaUnit = "m",
                              sampleSizeValue = NA,
                              sampleSizeUnit = NA,
                              footprintWKT = NA,
                              footprintSRS = NA,
                              isVegetationCoverReported = "false",
                              eventDate = tapply(
                                tick.occurrence$tck_fielddata$collectDate,
                                tick.occurrence$tck_fielddata$plotID,
                                function(x) paste(min(x, na.rm = TRUE), max(x, na.rm = TRUE), sep = "/")
                              )[NEON_plot_metadata$plotID],
                              eventTime = NA,
                              eventDurationValue = NA,
                              eventDurationUnit = NA,
                              eventType = "plot", #Q: This was 'Survey' in the mapping sheet
                              inventoryTypes = NA,
                              compilationTypes = NA,
                              compilationSourceTypes = NA,
                              samplingProtocol = "Drag sampling | Flagging", #Sampling protocol terms should be populated for every Event regardless of hierarchical level Q: I took this to mean all the terms described in 5.5. Methodology or sampling protocol (from here up to eventRemarks), or is this just for the four protocol fields?
                              protocolNames = "Drag sampling | Flagging",
                              protocolDescriptions = "Ticks are sampled at six plots per site every three to six weeks, depending on prior detections. Sampling follows the 160 m perimeter of each 40 × 40 m plot, using drag cloths (or flagging in dense vegetation). Ticks from all life stages are collected every few meters and preserved in ethanol, with bouts scheduled between green-up and dormancy. Samples are sent to external labs where they are sorted, counted and a subset of identified nymphs are tested for pathogens.",
                              protocolReferences = "Paull, S. 2022. TOS Protocol and Procedure: TCK – Tick and Tick-Borne Pathogen Sampling. NEON.DOC.014045. NEON (National Ecological Observatory Network). | Laboratory of Medical Zoology (LMZ). 2023. NEON Tick Pathogen Testing SOP, V4.01. University of Massachusetts, Amherst. | Beati, L. 2021. Tick Identification Instructions, USNTC Standard Operating Procedure (SOP). V3. Georgia Southern University, US National Tick Collection (USNTC).",
                              isAbsenceReported = "true", #Q: Should this be reported true at this level or only at the event visit level?
                              absentTaxa = NA,
                              isAbundanceReported = "true",
                              isAbundanceCapReported = "false",
                              abundanceCap = NA,
                              hasMaterialSamples = "true", #Q: Since we're not explicitly designating these as material samples, should this be false?
                              materialSampleTypes = "wholeOrganism",
                              hasVouchers = "true",
                              voucherInstitutions = "US National Tick Collection, Georgia Southern University",
                              isLeastSpecificTargetCategoryQuantityInclusive = "false",
                              dataGeneralizations = NA,
                              informationWithheld = NA, #TODO: double check this - I don't think any tick taxa are withheld but I'm not 100% certain
                              fieldNotes = NA,
                              eventRemarks = NA,
                              verbatimTargetScope = "ticks", # Recommended best practice is to populate scope terms every Event to which they apply.
                              targetTaxonomicScope = "?Ixodidae", #Q: Since we have pathogens in occurrences, I'm not sure how to designate those here - the ones tested don't have a mca - most are Bacteria but one is a piroplasm
                              excludedTaxonomicScope = NA,
                              isTaxonomicScopeFullyReported = "true",
                              taxonCompletenessReported = "reportedComplete", #Q: I'm not 100% sure I did these two correctly
                              taxonCompletenessProtocols = "Tick drag/flag methods are standardized to comprehensively detect all life stages of Ixodidae during the active season.", #Q: I'm not 100% sure I did these two correctly
                              hasNonTargetTaxa = "false",
                              areNonTargetTaxaFullyReported = NA,
                              nonTargetTaxa = NA,
                              identifiedBy = tapply(
                                tick.occurrence$tck_taxonomyProcessed$laboratoryName,
                                tick.occurrence$tck_taxonomyProcessed$plotID,
                                function(x) paste(sort(unique(x)), collapse = " | ") #Q: Some plots have had no ticks identified, so identifiedBy is NA - should this be changed to something more descriptive, like 'No ticks found'?
                              )[NEON_plot_metadata$plotID], #Q: I could change this to get the actual list of people who did the identifications, though this is not clean
                              targetLifeStageScope = "adult | nymph | larvae",
                              excludedLifeStageScope = "egg",
                              isLifeStageScopeFullyReported = "true",
                              targetDegreeOfEstablishmentScope = NA,
                              excludedDegreeOfEstablishmentScope = NA,
                              isDegreeOfEstablishmentScopeFullyReported = NA,
                              targetGrowthFormScope = NA,
                              excludedGrowthFormScope = NA,
                              isGrowthFormScopeFullyReported = NA,
                              hasNonTargetOrganisms = NA,
                              targetHabitatScope = NA,
                              excludedHabitatScope = NA,
                              samplingEffort = "160 m perimeter sampled per bout",
                              isSamplingEffortReported = "true",
                              samplingEffortProtocol = "Ticks are sampled at six plots per site, with bouts every 3 or 6 weeks depending on intensity. Each bout consists of dragging (or flagging if needed) along the full 160 m perimeter of each 40 × 40 m plot. Ticks of all life stages are collected at intervals and preserved in ethanol.",
                              samplingEffortValue = "6, 160",
                              samplingEffortUnit = "plots per bout, meters per plot circuit",
                              samplingPerformedBy = "NEON Field Staff" #Q: I could change this to get the list of actual people, though this has orcid and emails and is quite messy
                            )

```

### Plot Visit Events

Q: I didn't map sampleCondition since we originally had this going to materialEntityRemarks. Since we're not doing materialSample, is this where it should still go?

```{r}
NEON_plot_visit_metadata <- tick.occurrence$tck_fielddata

added_plot_visit_data <- data.frame(parentEventID = NEON_plot_visit_metadata$plotID,
                                    eventID = ifelse(is.na(NEON_plot_visit_metadata$sampleID), #eventID is the sampleID if a sample is taken, if not, we create one with the same format (plotID.date)
                                                      paste0(
                                                        NEON_plot_visit_metadata$plotID, ".",
                                                        format(as.Date(NEON_plot_visit_metadata$collectDate), "%Y%m%d")
                                                      ),
                                                      NEON_plot_visit_metadata$sampleID
                                                    ),
                                    siteNestingDescription = "One visit to a 40x40 m plot designated for tick sampling",
                                    siteCount = 1, #Q: I don't know if this should be here for this event
                                    fieldNumber = NA,
                                    verbatimSiteNames = NEON_plot_visit_metadata$plotID,
                                    habitat = NEON_plot_visit_metadata$nlcdClass,
                                    verbatimSiteDescriptions = NA,
                                    reportedWeather = NA,
                                    reportedExtremeConditions = ifelse(
                                      is.na(NEON_plot_visit_metadata$samplingImpractical) | NEON_plot_visit_metadata$samplingImpractical == "OK", NA,
                                      paste0(NEON_plot_visit_metadata$samplingImpractical, ", sampling impractical")
                                    ),
                                    locationID = NEON_plot_visit_metadata$namedLocation,
                                    countryCode = "US",
                                    decimalLatitude = NEON_plot_visit_metadata$decimalLatitude,
                                    decimalLongitude = NEON_plot_visit_metadata$decimalLongitude,
                                    coordinateUncertaintyInMeters = NEON_plot_visit_metadata$coordinateUncertainty,
                                    geodeticDatum = "WGS84",
                                    locality = NEON_plot_visit_metadata$plotID,
                                    minimumElevationinMeters = NEON_plot_visit_metadata$elevation - NEON_plot_visit_metadata$elevationUncertainty,
                                    maximumElevationinMeters = NEON_plot_visit_metadata$elevation + NEON_plot_visit_metadata$elevationUncertainty,
                                    verbatimElevation = paste(NEON_plot_visit_metadata$elevation, "m", sep=""),
                                    totalAreaSampledValue = NEON_plot_visit_metadata$totalSampledArea,
                                    totalAreaSampledUnit = ifelse(
                                      is.na(NEON_plot_visit_metadata$totalSampledArea), NA, "m"
                                    ),
                                    geospatialScopeAreaValue = NA,
                                    geospatialScopeAreaUnit = NA,
                                    sampleSizeValue = NA,
                                    sampleSizeUnit = NA,
                                    footprintWKT = NA,
                                    footprintSRS = NA,
                                    isVegetationCoverReported = "false",
                                    eventDate = paste(NEON_plot_visit_metadata$collectDate),
                                    eventTime = NA,
                                    eventDurationValue = NA,
                                    eventDurationUnit = NA,
                                    eventType = "plot visit", #Q: This was 'Site Visit' in the mapping sheet
                                    inventoryTypes = NA,
                                    compilationTypes = NA,
                                    compilationSourceTypes = NA,
                                    samplingProtocol = NEON_plot_visit_metadata$samplingMethod, #Sampling protocol terms should be populated for every Event regardless of hierarchical level Q: I took this to mean all the terms described in 5.5. Methodology or sampling protocol (from here up to eventRemarks), or is this just for the four protocol fields?
                                    protocolNames = NEON_plot_visit_metadata$samplingMethod,
                                    protocolDescriptions = "Ticks are sampled at six plots per site every three to six weeks, depending on prior detections. Sampling follows the 160 m perimeter of each 40 × 40 m plot, using drag cloths (or flagging in dense vegetation). Ticks from all life stages are collected every few meters and preserved in ethanol, with bouts scheduled between green-up and dormancy. Samples are sent to external labs where they are sorted, counted and a subset of identified nymphs are tested for pathogens.",
                                    protocolReferences = paste(NEON_plot_visit_metadata$samplingProtocolVersion, "Paull, S. 2022. TOS Protocol and Procedure: TCK – Tick and Tick-Borne Pathogen Sampling. NEON.DOC.014045. NEON (National Ecological Observatory Network). | Laboratory of Medical Zoology (LMZ). 2023. NEON Tick Pathogen Testing SOP, V4.01. University of Massachusetts, Amherst. | Beati, L. 2021. Tick Identification Instructions, USNTC Standard Operating Procedure (SOP). V3. Georgia Southern University, US National Tick Collection (USNTC).", sep=" | "),
                                    isAbsenceReported = ifelse(
                                      is.na(NEON_plot_visit_metadata$targetTaxaPresent),
                                      "false",
                                      "true"
                                    ), #Q: I only did this at the event level, not the occurrence level (i.e., there are no occurrences with occurrenceStatus = absent) - the guide says otherwise #Q: if targetTaxaPresent is NA (ie. there was no sampling done, should this be NA or false?) #Q: I think I'm getting lost on if this is an attribute of the survey event visit or the protocol
                                    absentTaxa = ifelse(
                                      NEON_plot_visit_metadata$targetTaxaPresent == "N", "?Ixodidae", NA #Q: Since we have pathogens in occurrences, I'm not sure how to designate those here - the ones tested don't have a mca - most are Bacteria but one is a piroplasm  #Q: If there are no absences (i.e., there were ticks found) should this be something besides NA? Should this be a list of species?
                                    ),
                                    isAbundanceReported = ifelse(
                                      is.na(NEON_plot_visit_metadata$targetTaxaPresent),
                                      "false",
                                      "true"
                                    ), #Q: I think I'm getting lost on if this is an attribute of the survey event visit or the protocol #Q: If no survey occurred should this be something besides NA?
                                    isAbundanceCapReported = "false",
                                    abundanceCap = NA,
                                    hasMaterialSamples = "true", #Q: Since we're not explicitly designating these as material samples, should this be false?
                                    materialSampleTypes = "wholeOrganism",
                                    hasVouchers = "true",
                                    voucherInstitutions = "US National Tick Collection, Georgia Southern University",
                                    isLeastSpecificTargetCategoryQuantityInclusive = "false",
                                    dataGeneralizations = NA,
                                    informationWithheld = NA, #TODO: double check this - I don't think any tick taxa are withheld but I'm not 100% certain
                                    fieldNotes = NA,
                                    eventRemarks = NEON_plot_visit_metadata$remarks,
                                    verbatimTargetScope = "ticks", # Recommended best practice is to populate scope terms every Event to which they apply.
                                    targetTaxonomicScope = "?Ixodidae", #Q: Since we have pathogens in occurrences, I'm not sure how to designate those here - the ones tested don't have a mca - most are Bacteria but one is a piroplasm
                                    excludedTaxonomicScope = NA,
                                    isTaxonomicScopeFullyReported = "true",
                                    taxonCompletenessReported = "reportedComplete", #Q: I'm not 100% sure I did these two correctly
                                    taxonCompletenessProtocols = "Tick drag/flag methods are standardized to comprehensively detect all life stages of Ixodidae during the active season.", #Q: I'm not 100% sure I did these two correctly
                                    hasNonTargetTaxa = "false",
                                    areNonTargetTaxaFullyReported = NA,
                                    nonTargetTaxa = NA,
                                    identifiedBy = tick.occurrence$tck_fielddata %>% #Q: Some plots have had no ticks identified, so identifiedBy is NA - should this be changed to something more descriptive, like 'No ticks found'?
                                      left_join(
                                        tick.occurrence$tck_taxonomyProcessed %>%
                                          group_by(sampleID) %>%
                                          summarise(
                                            identifiedBy = paste(sort(unique(identifiedBy)), collapse = " | "),
                                            .groups = "drop"
                                          ),
                                        by = "sampleID"
                                      ) %>%
                                      pull(identifiedBy),
                                    targetLifeStageScope = "adult | nymph | larvae",
                                    excludedLifeStageScope = "egg",
                                    isLifeStageScopeFullyReported = "true",
                                    targetDegreeOfEstablishmentScope = NA,
                                    excludedDegreeOfEstablishmentScope = NA,
                                    isDegreeOfEstablishmentScopeFullyReported = NA,
                                    targetGrowthFormScope = NA,
                                    excludedGrowthFormScope = NA,
                                    isGrowthFormScopeFullyReported = NA,
                                    hasNonTargetOrganisms = NA,
                                    targetHabitatScope = NA,
                                    excludedHabitatScope = NA,
                                    samplingEffort = "160 m perimeter sampled per bout",
                                    isSamplingEffortReported = "true",
                                    samplingEffortProtocol = "Ticks are sampled at six plots per site, with bouts every 3 or 6 weeks depending on intensity. Each bout consists of dragging (or flagging if needed) along the full 160 m perimeter of each 40 × 40 m plot. Ticks of all life stages are collected at intervals and preserved in ethanol.",
                                    samplingEffortValue = "6, 160",
                                    samplingEffortUnit = "plots per bout, meters per plot circuit",
                                    samplingPerformedBy = NEON_plot_visit_metadata$measuredBy
                                  )

```

### Merge Event Tables

```{r}
event_data <- bind_rows(
  NEON_project_event,
  added_domain_data,
  added_site_data,
  added_plot_data,
  added_plot_visit_data
)
rownames(event_data) <- NULL
```


## Map Occurrence Data

### Tick Occurrences

```{r}
NEON_tick_occurrence_data <- tick.occurrence$tck_taxonomyProcessed

added_tick_occurrence_data <- data.frame(eventID = NEON_tick_occurrence_data$sampleID,
                                         occurrenceID = NEON_tick_occurrence_data$subsampleID,
                                         basisOfRecord = "PreservedSpecimen",
                                         scientificName = NEON_tick_occurrence_data$scientificName,
                                         scientificNameAuthorship = NEON_tick_occurrence_data$scientificNameAuthorship,
                                         taxonRank = NEON_tick_occurrence_data$taxonRank,
                                         kingdom = "Animalia",
                                         family = NEON_tick_occurrence_data$family,
                                         subfamily = NEON_tick_occurrence_data$subfamily,
                                         tribe = NEON_tick_occurrence_data$tribe,
                                         subtribe = NEON_tick_occurrence_data$subtribe,
                                         genus = NEON_tick_occurrence_data$genus,
                                         subgenus = NEON_tick_occurrence_data$subgenus,
                                         specificEpithet = NEON_tick_occurrence_data$specificEpithet,
                                         infraspecificEpithet = NEON_tick_occurrence_data$infraspecificEpithet,
                                         identificationQualifier = NEON_tick_occurrence_data$identificationQualifier,
                                         identificationReferences = NEON_tick_occurrence_data$identificationReferences,
                                         identificationRemarks = NEON_tick_occurrence_data$identificationProtocolVersion,
                                         occurrenceStatus = "present",
                                         organismQuantity = NA, #individualCount is used instead
                                         organismQuantityType = NA, #individualCount is used instead
                                         individualCount = NEON_tick_occurrence_data$individualCount,
                                         recordedByID = NA, #This information is in the plot visit event
                                         vernacularName = c("Dermacentor variabilis" = "American dog tick",
                                                            "Ixodes scapularis" = "Blacklegged tick / Deer tick",
                                                            "Haemaphysalis leporispalustris" = "Rabbit tick",
                                                            "Amblyomma americanum" = "Lone star tick",
                                                            "Haemaphysalis longicornis" = "Asian longhorned tick",
                                                            "Ixodes muris" = "Mouse tick",
                                                            "Ixodes dentatus" = "Rabbit tick",
                                                            "Amblyomma maculatum" = "Gulf Coast tick",
                                                            "Ixodes marxi" = "Squirrel tick",
                                                            "Ixodes brunneus" = "Bird tick",
                                                            "Dermacentor andersoni" = "Rocky Mountain wood tick",
                                                            "Dermacentor parumapertus" = "Desert wood tick",
                                                            "Ixodes pacificus" = "Western blacklegged tick",
                                                            "Ixodes angustus" = "Mouse tick",
                                                            "Dermacentor occidentalis" = "Pacific Coast tick"
                                                          )[NEON_tick_occurrence_data$scientificName],
                                         sex = ifelse(NEON_tick_occurrence_data$sexOrAge %in% c("Male", "Female"),
                                                       NEON_tick_occurrence_data$sexOrAge,
                                                       NA),
                                         lifeStage = ifelse(NEON_tick_occurrence_data$sexOrAge %in% c("Male", "Female"),
                                                             "Adult",
                                                             NEON_tick_occurrence_data$sexOrAge),
                                         establishmentMeans = NA,
                                         degreeOfEstablishment = NA,
                                         pathway = NA,
                                         vitality = "alive",
                                         preparations = NEON_tick_occurrence_data$archiveMedium,
                                         institutionID = NEON_tick_occurrence_data$archiveFacilityID,
                                         identifiedBy = paste(NEON_tick_occurrence_data$identifiedBy, NEON_tick_occurrence_data$laboratoryName, sep=", "),
                                         occurrenceRemarks = NEON_tick_occurrence_data$remarks
                               )

```

### Tick Pathogen Occurrences
```{r}
NEON_pathogen_occurrence_data <- tick.pathogen$tck_pathogen %>%
  filter(
    !is.na(testResult) &
    !is.na(testPathogenName) &
      testPathogenName != "HardTick DNA Quality" & #Remove DNA Quality test
      !grepl("^(Ixodes|Dermacentor|Amblyomma|Haemaphysalis)", testPathogenName) #Remove molecular identification tests
  )

added_pathogen_occurrence_data <- data.frame(eventID = sapply(strsplit(NEON_pathogen_occurrence_data$subsampleID, "\\."), 
                                                              function(x) paste(x[1:2], collapse = ".")),
                                             occurrenceID = NEON_pathogen_occurrence_data$uid,
                                             basisOfRecord = "LaboratoryObservation",  #Q: I'm not sure what to put here
                                             scientificName = NEON_pathogen_occurrence_data$testPathogenName,
                                             taxonRank = NA,
                                             kingdom = c("Borrelia burgdorferi" = "Bacteria",
                                                         "Borrelia miyamotoi" = "Bacteria",
                                                         "Anaplasma phagocytophilum" = "Bacteria",
                                                         "Rickettsia rickettsii" = "Bacteria",
                                                         "Ehrlichia chaffeensis" = "Bacteria",
                                                         "Borrelia lonestari" = "Bacteria",
                                                         "Babesia microti" = "Protista",
                                                         "Ehrlichia ewingii" = "Bacteria",
                                                         "Borrelia burgdorferi sensu lato" = "Bacteria",
                                                         "Ehrlichia muris-like" = "Bacteria",
                                                         "Borrelia mayonii" = "Bacteria",
                                                         "Francisella tularensis" = "Bacteria",
                                                         "Borrelia sp." = "Bacteria",
                                                         "Rickettsia parkeri" = "Bacteria",
                                                         "Rickettsia philipii" = "Bacteria"
                                                         )[NEON_pathogen_occurrence_data$testPathogenName],
                                             occurrenceStatus = ifelse(NEON_pathogen_occurrence_data$testResult == "Negative", "absent",
                                                                       ifelse(NEON_pathogen_occurrence_data$testResult == "Positive", "present", NA)
                                                                      ),
                                             organismQuantity = NA,
                                             organismQuantityType = NA,
                                             individualCount = NA, #Recommend not sharing this until we find out specifically what this is actually reporting. Note: this is always 1.
                                             recordedByID = NA, #This information is in the plot visit event
                                             vernacularName = c("Borrelia burgdorferi" = "Lyme disease agent",
                                                                "Borrelia miyamotoi" = "Borrelia miyamotoi disease (BMD) agent",
                                                                "Anaplasma phagocytophilum" = "Human granulocytic anaplasmosis (HGA) agent",
                                                                "Rickettsia rickettsii" = "Rocky Mountain spotted fever agent",
                                                                "Ehrlichia chaffeensis" = "Human monocytic ehrlichiosis agent",
                                                                "Borrelia lonestari" = "Southern tick-associated rash illness (STARI) agent",
                                                                "Babesia microti" = "Babesiosis parasite",
                                                                "Ehrlichia ewingii"  = "Human ehrlichiosis agent",
                                                                "Borrelia burgdorferi sensu lato" = "Lyme disease agent",
                                                                "Ehrlichia muris-like" = "Human ehrlichiosis agent",
                                                                "Borrelia mayonii" = "Lyme disease agent",
                                                                "Francisella tularensis" = "Tularemia agent",
                                                                "Borrelia sp." = "Lyme disease or relapsing fever agent",
                                                                "Rickettsia parkeri" = "Rickettsia parkeri rickettsiosis agent",
                                                                "Rickettsia philipii" = "Pacific coast tick fever agent"
                                                               )[NEON_pathogen_occurrence_data$testPathogenName],
                                             sex = NA,
                                             lifeStage = NA,
                                             establishmentMeans = NA,
                                             degreeOfEstablishment = NA,
                                             pathway = NA,
                                             vitality = NA,
                                             identificationRemarks = paste(NEON_pathogen_occurrence_data$testProtocolVersion,
                                                                            ifelse(is.na(NEON_pathogen_occurrence_data$remarks), "", 
                                                                                   paste("; ", NEON_pathogen_occurrence_data$remarks))
                                                                          ),
                                             identifiedBy = paste(NEON_pathogen_occurrence_data$testedBy, NEON_pathogen_occurrence_data$laboratoryName, sep=", "),
                                             dateIdentified = NEON_pathogen_occurrence_data$testedDate,
                                             materialSampleID = NEON_pathogen_occurrence_data$testingID, #Q: Does this need to be remapped or removed since we aren't doing materialSample?
                                             materialEntityRemarks = "materialSampleID corresponds to the NEON testingID, which uniquely identifies the individual tick nymph(s) tested."
                                   )
```

### Merge Occurrence Tables
```{r}
occurrence_data <- bind_rows(added_tick_occurrence_data,
                             added_pathogen_occurrence_data)
rownames(occurrence_data) <- NULL
```


## Map Resource Relationship Data
```{r}
resource_relationship_data <- data.frame(resourceID = NEON_pathogen_occurrence_data$uid, #occurrenceID (pathogen)
                                         relatedResourceID = NEON_pathogen_occurrence_data$subsampleID, #occurrenceID (tick)
                                         relationshipOfResource = "pathogen of"
)
```

## Map Extended Measurement or Fact Data

### Plot Type
```{r}
emof_plot_data <- data.frame(eventID = NEON_plot_metadata$plotID,
                             occurrenceID = NA, # This is just for readability of the final table
                             measurementType = "Plot Type",
                             measurementValue = NEON_plot_metadata$plotType
)
```

### Counts
```{r}
#Q: I removed records where these are NA, should I keep those in (i.e., and have a record for each event regardless)?
emof_counts_adult_data <- data.frame( #only create a row if adultCount is not NA
  eventID = ifelse( 
    is.na(NEON_plot_visit_metadata$sampleID), #eventID is the sampleID if a sample is taken, if not, we create one with the same format (plotID.date)
    paste0(NEON_plot_visit_metadata$plotID, ".",
           format(as.Date(NEON_plot_visit_metadata$collectDate), "%Y%m%d")
    ),
    NEON_plot_visit_metadata$sampleID
  )[!is.na(NEON_plot_visit_metadata$adultCount)],
  occurrenceID = NA, # This is just for readability of the final table
  measurementType = "Adult Count",
  measurementValue = as.character(NEON_plot_visit_metadata$adultCount[!is.na(NEON_plot_visit_metadata$adultCount)])
)

emof_counts_nymph_data <- data.frame( #only create a row if nymphCount is not NA
  eventID = ifelse( 
    is.na(NEON_plot_visit_metadata$sampleID), #eventID is the sampleID if a sample is taken, if not, we create one with the same format (plotID.date)
    paste0(NEON_plot_visit_metadata$plotID, ".",
           format(as.Date(NEON_plot_visit_metadata$collectDate), "%Y%m%d")
    ),
    NEON_plot_visit_metadata$sampleID
  )[!is.na(NEON_plot_visit_metadata$nymphCount)],
  occurrenceID = NA, # This is just for readability of the final table
  measurementType = "Nymph Count",
  measurementValue = as.character(NEON_plot_visit_metadata$nymphCount[!is.na(NEON_plot_visit_metadata$nymphCount)])
)

emof_counts_larva_data <- data.frame( #only create a row if larvaCount is not NA
  eventID = ifelse( 
    is.na(NEON_plot_visit_metadata$sampleID), #eventID is the sampleID if a sample is taken, if not, we create one with the same format (plotID.date)
    paste0(NEON_plot_visit_metadata$plotID, ".",
           format(as.Date(NEON_plot_visit_metadata$collectDate), "%Y%m%d")
    ),
    NEON_plot_visit_metadata$sampleID
  )[!is.na(NEON_plot_visit_metadata$larvaCount)],
  occurrenceID = NA, # This is just for readability of the final table
  measurementType = "Larva Count",
  measurementValue = as.character(NEON_plot_visit_metadata$larvaCount[!is.na(NEON_plot_visit_metadata$larvaCount)])
)
```

### Sample Code (Barcode)
```{r}
emof_sample_code_data <- data.frame( #only create a row if sampleCount is not NA
  eventID = ifelse(
    is.na(NEON_plot_visit_metadata$sampleID), #eventID is the sampleID if a sample is taken, if not, we create one with the same format (plotID.date)
    paste0(NEON_plot_visit_metadata$plotID, ".",
           format(as.Date(NEON_plot_visit_metadata$collectDate), "%Y%m%d")
    ),
    NEON_plot_visit_metadata$sampleID
  )[!is.na(NEON_plot_visit_metadata$sampleCode)],
  occurrenceID = NA, # This is just for readability of the final table
  measurementType = "Sample Barcode",
  measurementValue = NEON_plot_visit_metadata$sampleCode[!is.na(NEON_plot_visit_metadata$sampleCode)]
)
```

### Tick Sample Condition
```{r}
emof_tick_sample_condition_data <- data.frame(eventID = NEON_tick_occurrence_data$sampleID,
                                              occurrenceID = NEON_tick_occurrence_data$subsampleID,
                                              measurementType = "Tick Sample Condition",
                                              measurementValue = NEON_tick_occurrence_data$sampleCondition
)
```

### Deprecated Vial ID
```{r}
emof_deprecatedVialID_data <- data.frame(eventID = NEON_tick_occurrence_data$sampleID,
                                         occurrenceID = NEON_tick_occurrence_data$subsampleID,
                                         measurementType = "Deprecated Vial ID",
                                         measurementValue = NEON_tick_occurrence_data$deprecatedVialID
)
```

### Tick Pathogen Sample Condition
```{r}
emof_pathogen_sample_condition_data <- data.frame(eventID = sapply(strsplit(NEON_pathogen_occurrence_data$subsampleID, "\\."), 
                                                                   function(x) paste(x[1:2], collapse = ".")),
                                                  occurrenceID = NEON_pathogen_occurrence_data$uid,
                                                  measurementType = "Tick Pathogen Sample Condition",
                                                  measurementValue = NEON_pathogen_occurrence_data$sampleCondition
)
```

### CQ (Quantification Cycle or Threshold Number) Value
```{r}
emof_cqValue_data <- data.frame(
  eventID = sapply(strsplit(NEON_pathogen_occurrence_data$subsampleID[!is.na(NEON_pathogen_occurrence_data$cqValue)], "\\."),
                   function(x) paste(x[1:2], collapse = ".")),
  occurrenceID = NEON_pathogen_occurrence_data$uid[!is.na(NEON_pathogen_occurrence_data$cqValue)],
  measurementType = "CQ (Quantification Cycle or Threshold Number) Value",
  measurementValue = as.character(NEON_pathogen_occurrence_data$cqValue[!is.na(NEON_pathogen_occurrence_data$cqValue)])
)
```

### DNA Sample ID
```{r}
emof_dnaSampleID_data <- data.frame(
  eventID = sapply(strsplit(NEON_pathogen_occurrence_data$subsampleID[!is.na(NEON_pathogen_occurrence_data$dnaSampleID)], "\\."),
                   function(x) paste(x[1:2], collapse = ".")),
  occurrenceID = NEON_pathogen_occurrence_data$uid[!is.na(NEON_pathogen_occurrence_data$dnaSampleID)],
  measurementType = "Identifier for DNA sample",
  measurementValue = NEON_pathogen_occurrence_data$dnaSampleID[!is.na(NEON_pathogen_occurrence_data$dnaSampleID)]
)
```

### DNA Sample Code
```{r}
emof_dnaSampleCode_data <- data.frame(
  eventID = sapply(strsplit(NEON_pathogen_occurrence_data$subsampleID[!is.na(NEON_pathogen_occurrence_data$dnaSampleCode)], "\\."),
                   function(x) paste(x[1:2], collapse = ".")),
  occurrenceID = NEON_pathogen_occurrence_data$uid[!is.na(NEON_pathogen_occurrence_data$dnaSampleCode)],
  measurementType = "Barcode of a DNA sample",
  measurementValue = NEON_pathogen_occurrence_data$dnaSampleCode[!is.na(NEON_pathogen_occurrence_data$dnaSampleCode)]
)
```

### Merge Extended Measurement or Fact Tables
```{r}
emof_data <- bind_rows(emof_plot_data,
                       emof_counts_adult_data,
                       emof_counts_nymph_data,
                       emof_counts_larva_data,
                       emof_sample_code_data,
                       emof_tick_sample_condition_data,
                       emof_deprecatedVialID_data,
                       emof_pathogen_sample_condition_data,
                       emof_cqValue_data,
                       emof_dnaSampleID_data,
                       emof_dnaSampleCode_data
                       )
rownames(emof_data) <- NULL
```

# Save Data
```{r}
write.csv(event_data, "outputs/event.csv", row.names = FALSE)
write.csv(occurrence_data, "outputs/occurrence.csv", row.names = FALSE)
write.csv(resource_relationship_data, "outputs/resourceRelationship.csv", row.names = FALSE)
write.csv(emof_data, "outputs/extendedMeasurementOrFact.csv", row.names = FALSE)
```
